Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-12-12 demystify.tex
updates

\section{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://www.gmu.edu/international/english-language-requirements}{these}, you will need to take standardized English tests. On one hand you will need to show some level of English proficiency, but on the other hand, you do not need to have very good scores in these tests (many adcom members themselves were once international students and struggled with English).
You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}).
Just as with grades (\autoref{chap:your-school}) and GRE, having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.

\begin{commentbox}[Vu:]
  Here is the minimum requirements at Mason. 
  Being above this might not mean much, but below is a \red{red flag}.
  \begin{itemize}
    \item GPA: $\ge 3.0$ in your undergrad (but we also consider the rank/prestige of your school)
    \item GRE: not required 
    % \item but if you want to use it, then we expect a total (V+Q) of $\ge 311$ (with a $\ge 157$ Q) and A $\ge 3.0-3.5$.
    \item \href{https://www.gmu.edu/international/english-language-requirements}{English proficiency requirements} (one of the below)
          \begin{itemize}
            \item TOEFL: 80 OR
            \item IELTS: $\ge 6.5$ OR
            \item DuoLingo Graduate English: $\ge 120$ OR
            \item Pearson Test of Academic English: $\ge 67$
          \end{itemize}
  \end{itemize}
\end{commentbox}

@@

Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.  For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.

@@

  \item Objective: an MS is typically to prepare you for \emph{industry}, while a PhD is to prepare you for research and academia. Some MS has thesis option but the research is much lighter compared to a PhD.

@@

You typically need to submit the similar materials as in a PhD application, e.g., transcripts, LoRs, SOP, and test scores.  However, research experience is not required, and LoRs can be from anyone who can speak about your academic or working abilities. Moreover, your SOP should focus on your academic and work background, why you want to get an MS, and how the program fits your goals. As with PhDs, GREs are often not required. Some MS programs do not even require LoRs or SOPs.

@@

\paragraph{Test Scores} GREs are often not required for MS applications.  However, some schools might still require them, and you should check with the school you're applying to.  If you have a low GPA, a high GRE score can help offset that. As with PhD, English proficiency tests (TOEFL/IELTS) are required for international students.

>>>>

\section{English Tests (IELTS, TOEFL)}\label{sec:english-tests}

Unless your degrees are from the US or certain countries such as \href{https://www.gmu.edu/international/english-language-requirements}{these}, you will need to take standardized English tests. On one hand you will need to show some level of English proficiency, but on the other hand, you do not need to have very good scores in these tests (many adcom members themselves were once international students and struggled with English).
You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}).
Just as with grades (\autoref{chap:your-school}) and GRE, having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.

\begin{commentbox}[Vu:]
  Here is the minimum requirements at Mason. 
  Being above this might not mean much, but below is a \red{red flag}.
  \begin{itemize}
    \item GPA: $\ge 3.0$ in your undergrad (but we also consider the rank/prestige of your school)
    \item GRE: not required 
    % \item but if you want to use it, then we expect a total (V+Q) of $\ge 311$ (with a $\ge 157$ Q) and A $\ge 3.0-3.5$.
    \item \href{https://www.gmu.edu/international/english-language-requirements}{English proficiency requirements} (one of the below)
          \begin{itemize}
            \item TOEFL: 80 OR
            \item IELTS: $\ge 6.5$ OR
            \item DuoLingo Graduate English: $\ge 120$ OR
            \item Pearson Test of Academic English: $\ge 67$
          \end{itemize}
  \end{itemize}
\end{commentbox}
% Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.  For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.

@@

@@

  \item Objective: an MS is typically to prepare you for \emph{industry}, while a PhD is to prepare you for research and academia. Some MS has thesis option but in general research is not a focus in MS programs.

@@

You will submit similar materials as in a PhD application, e.g., transcripts, LoRs, SOP, and test scores.  However, research experience is not required, and LoRs can be from anyone who can speak about your academic or working abilities. Moreover, your SOP should focus on your academic and work background, why you want to get an MS, and how the program fits your goals. As with PhDs, GREs are often not required. Some MS programs do not even require LoRs or SOPs.

@@

\paragraph{Test Scores} GREs are often not required for MS applications.  However, some schools might still require them, and you should check with the school you're applying to.  If you have a low GPA, a high GRE score can help offset that. As with PhD, English proficiency tests (TOEFL/IELTS) are required for international students (but just passing the mimimum requirement is often enough \autoref{sec:english-tests}). Note that just as with PhD, Englih tests are are waived if you did your undergrad at a US university.
